---
title: "pacman"
output: html_document
vignette: >
  %\VignetteIndexEntry{pacman}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 5
)
```

```{r setup}
library(grid)
library(nara)
```

## Pacman

This vignette shows how to build a Pacman game board and animate the 
sprites within to produce a system which displays the movement in realtime.


<video controls>
  <source src="./image/pacman-attractb.mp4" type="video/mp4">
  Your browser does not support the video tag.
</video> 


#### To turn this into an actual game

This vignette is just a demo of the graphical framework for a game using the
`{nara}` package.  It is not meant to be a playable game.

Many things be needed to make this into an actual game.  Listed in 
a rough order of importance:

1. User control with `{eventloop}`
2. pacman/ghost collision detection. A check if they're both simultaneously
   at the same coordinates would be sufficient for a first draft
3. Scoring
4. End condition check
5. More levels
6. Actual ghost behaviour (which is more complex than constant-velocity random 
   movement)
7. Power dots
8. Sound

## Game Board

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#' Helper function: Reverse the characters in a string
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
str_rev <- function(x) {
  paste(rev(strsplit(x, '')[[1]]), collapse="")
}

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Helper function: Swap 2 characters in a string.  
# Dodgy implementation. Good enough for now.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
str_swap <- function(x, s1, s2) {
  x <- gsub(s1 , 'x', x)
  x <- gsub(s2 ,  s1, x)
  x <- gsub('x',  s2, x)
  x
}

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Define the left half of the board
# Board is symmetrical.
# Full board = 31 x 28
#
# Define "obstruction" pieces.  e.g. "1" represents a rounded quarter-circle
# corner in the top-left of a square
#
#   1 2 3
#   4 5 6
#   7 8 9
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
left <- c(
"12222222222223", 
"4............6", 
"4.1223.12223.6", 
"4.6554.45556.6", 
"4.7889.78889.7", 
"4.............", 
"4.1223.13.1222", 
"4.7889.46.7883", 
"4......46....4", 
"788883.47883.4", 
"555556.41229.7", 
"555556.46.....", 
"555556.46.1222",
"555556.79.4555",
"555556....4555", # Middle
"555556.13.4555", 
"555556.46.7888", 
"555556.46.....",
"555556.46.1222",
"122229.79.7883",
"4............6",
"4.1223.12223.6",
"4.7834.78889.7",
"4...46........",
"783.46.13.1222",
"129.79.46.7883",
"4......46....6",
"4.1222297223.6",
"4.7888888889.7",
"4.............",
"78888888888888"
)


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Mirror the left of the board and update the tile references
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
right <- purrr::map_chr(left, str_rev)
right <- str_swap(right, '1', '3')
right <- str_swap(right, '4', '6')
right <- str_swap(right, '7', '9')

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Rearracnge board into a 31*28 character matrix
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
board <- purrr::map2_chr(left, right, ~paste0(.x, .y, collapse = ""))
board <- unlist(stringr::str_split(board, ''))
board <- matrix(board, nrow = 31, ncol = 28, byrow = TRUE)
```

## Board pathways

```{r echo = FALSE}
mat <- board
mat[] <- ifelse(board == '.', 'black', 'white')
# mat

grid.raster(mat, interpolate = FALSE)
```


## Maze parts

```{r fig.width = 6, fig.height=4}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#' Load the maze parts
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
im <- png::readPNG("image/game-maze.png")
dim(im)
grid.raster(im)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Parse out the pacman sprites
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
s <- vector('list', 9)
row <- 1; col <- 7; s[[1]] <- im[28 + row*9 + 0:7, 226 + col*9 + 0:7,]
row <- 1; col <- 5; s[[2]] <- im[28 + row*9 + 0:7, 226 + col*9 + 0:7,]
row <- 1; col <- 6; s[[3]] <- im[28 + row*9 + 0:7, 226 + col*9 + 0:7,]

row <- 1; col <- 8; s[[4]] <- im[28 + row*9 + 0:7, 226 + col*9 + 0:7,]
row <- 2; col <-12; s[[5]] <- im[28 + row*9 + 0:7, 226 + col*9 + 0:7,]
row <- 1; col <- 9; s[[6]] <- im[28 + row*9 + 0:7, 226 + col*9 + 0:7,]

row <- 1; col <-11; s[[7]] <- im[28 + row*9 + 0:7, 226 + col*9 + 0:7,]
row <- 1; col <- 5; s[[8]] <- im[28 + row*9 + 0:7, 226 + col*9 + 0:7,]
row <- 1; col <-10; s[[9]] <- im[28 + row*9 + 0:7, 226 + col*9 + 0:7,]

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Convert from array to nativeraster
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
s <- lapply(s, nara::array_to_nr)
```

#### 3x3 grid of the maze pieces

```{r echo=FALSE}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create a montage 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
nr <- nr_new(3*8 + 3, 3*8 + 3, fill = 'white')
nr_blit(nr, s[[1]], 1 + 0 * 8 + 0, 1 + 2 * 8 + 2)
nr_blit(nr, s[[2]], 1 + 1 * 8 + 1, 1 + 2 * 8 + 2)
nr_blit(nr, s[[3]], 1 + 2 * 8 + 2, 1 + 2 * 8 + 2)

nr_blit(nr, s[[4]], 1 + 0 * 8 + 0, 1 + 1 * 8 + 1)
nr_blit(nr, s[[5]], 1 + 1 * 8 + 1, 1 + 1 * 8 + 1)
nr_blit(nr, s[[6]], 1 + 2 * 8 + 2, 1 + 1 * 8 + 1)

nr_blit(nr, s[[7]], 1 + 0 * 8 + 0, 1 + 0 * 8 + 0)
nr_blit(nr, s[[8]], 1 + 1 * 8 + 1, 1 + 0 * 8 + 0)
nr_blit(nr, s[[9]], 1 + 2 * 8 + 2, 1 + 0 * 8 + 0)

grid.newpage()
grid.raster(nr, interpolate = FALSE)
```


## Paste maze pieces to make board

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create a blank native raster for the board
# 31 squares high. 28 squares wide
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
blank_board_nr <- nr_new(width = 28 * 8, height = 31 * 8, fill = 'black')

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Copy the appropriate maze piece into the board
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
for (row in 1:31) {
  for (col in 1:28) {
    val <- board[31 + 1 - row, col]
    if (val != '.') {
      idx <- as.integer(val)
      nr_blit(blank_board_nr, s[[idx]], (col - 1) * 8 + 1, (row - 1) * 8 + 1)
    }
  }
}

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Draw the blank board
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
grid.newpage()
grid.raster(blank_board_nr, interpolate = FALSE)
```

## Define the starting dots

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
dots <- arrayInd(which(board == '.'), dim(board))
dots[,1] <- 32 - dots[,1] # flip y
dots <- as.data.frame(dots)
names(dots) <- c('y', 'x')


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create a 2x2 pixel nativeraster to represented the dot onscreen
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
dot_mat <- matrix(rep('white', 4), 2, 2)
dot_nr  <- nara::raster_to_nr(dot_mat)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Test drawing the dots over the blank board
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
nr <- nr_duplicate(blank_board_nr)  
nr_blit(nr, dot_nr, (dots$x - 0.5) * 8, (dots$y - 0.5) * 8)

grid.newpage()
grid.raster(nr, interpolate = FALSE)
```


#### Extracting the character sprites

```{r fig.width = 9, fig.height=4}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Load the spritemap for pacman and the ghosts
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
spritemap <- png::readPNG("image/game-sprites.png")
dim(spritemap)
grid.raster(spritemap)
```


```{r  fig.width=2, fig.height=2}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#' Extract a sprite from the spritemap
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
extract <- function(row, col) {
  sprite <- spritemap[1 + (row-1)*16 + 0:15, 457 + (col-1)*16 + 0:15,]
  
  alpha <- sprite[,,1] == 1 | sprite[,,2] == 1 | sprite[,,3] == 1 
  alpha[] <- as.numeric(alpha)
  
  new <- c(sprite, alpha)
  d <- dim(sprite)
  d[3] <- 4
  dim(new) <- d
  
  nara::array_to_nr(new)
}


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Extract pacman sprites
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
pacman <- list(
  right = list(extract(1, 1), extract(1, 2), extract(1, 3), extract(1, 2)),
  left  = list(extract(2, 1), extract(2, 2), extract(1, 3), extract(2, 2)),
  up    = list(extract(3, 1), extract(3, 2), extract(1, 3), extract(3, 2)),
  down  = list(extract(4, 1), extract(4, 2), extract(1, 3), extract(4, 2))
)


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Extract ghost sprites
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
ghost1 <- list(
  right = list(extract(5, 1), extract(5, 2)),
  left  = list(extract(5, 3), extract(5, 4)),
  up    = list(extract(5, 5), extract(5, 6)),
  down  = list(extract(5, 7), extract(5, 8))
)

ghost2 <- list(
  right = list(extract(6, 1), extract(6, 2)),
  left  = list(extract(6, 3), extract(6, 4)),
  up    = list(extract(6, 5), extract(6, 6)),
  down  = list(extract(6, 7), extract(6, 8))
)

ghost3 <- list(
  right = list(extract(7, 1), extract(7, 2)),
  left  = list(extract(7, 3), extract(7, 4)),
  up    = list(extract(7, 5), extract(7, 6)),
  down  = list(extract(7, 7), extract(7, 8))
)

ghost4 <- list(
  right = list(extract(8, 1), extract(8, 2)),
  left  = list(extract(8, 3), extract(8, 4)),
  up    = list(extract(8, 5), extract(8, 6)),
  down  = list(extract(8, 7), extract(8, 8))
)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Combine all ghosts
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
ghost <- list(
  ghost1, ghost2, ghost3, ghost4
)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Test plot of a single sprite
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
grid.raster(ghost4$right[[1]], interpolate = FALSE)
```

## Sprite rendering demo

Generate a realtime animation of pacman and ghosts running across the screen.

```{r eval=FALSE}
library(grid)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Open a a double-buffered x11() device.
# - turn antialiasing off (don't need it for pixel rendering)
# - inibit storaget of a displaylist
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
x11(type = 'dbcairo', antialias = 'none')
dev.control('inhibit')

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# In-memory rendering buffer
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
nr <- nr_new(100, 30, fill = 'black')


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Loop and display
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
for (i in c((-20):180, (-20):180)) {
  nr_fill(nr, 'black')
  nr_blit(nr, pacman$right[[(bitwShiftR(i, 1) %% 4) + 1]], i     , 5)
  nr_blit(nr, ghost1$right[[(bitwShiftR(i, 1) %% 2) + 1]], i - 20, 5)
  nr_blit(nr, ghost2$right[[(bitwShiftR(i, 1) %% 2) + 1]], i - 40, 5)
  nr_blit(nr, ghost3$right[[(bitwShiftR(i, 1) %% 2) + 1]], i - 60, 5)
  nr_blit(nr, ghost4$right[[(bitwShiftR(i, 1) %% 2) + 1]], i - 80, 5)
  
  
  if (i < 20 + 0 * 11) nr_text(nr, "#", 20 + 0 * 11, 9, colour = 'white', fontsize = 11)
  if (i < 20 + 1 * 11) nr_text(nr, "R", 20 + 1 * 11, 9, colour = 'white', fontsize = 11)
  if (i < 20 + 2 * 11) nr_text(nr, "s", 20 + 2 * 11, 9, colour = 'white', fontsize = 11)
  if (i < 20 + 3 * 11) nr_text(nr, "t", 20 + 3 * 11, 9, colour = 'white', fontsize = 11)
  if (i < 20 + 4 * 11) nr_text(nr, "a", 20 + 4 * 11, 9, colour = 'white', fontsize = 11)
  if (i < 20 + 5 * 11) nr_text(nr, "t", 20 + 5 * 11, 9, colour = 'white', fontsize = 11)
  if (i < 20 + 6 * 11) nr_text(nr, "s", 20 + 6 * 11, 9, colour = 'white', fontsize = 11)
  
  dev.hold()
  grid.raster(nr, interpolate = FALSE)
  dev.flush()
  Sys.sleep(0.02)
}
```


<video controls>
  <source src="./image/pacman2b.mp4" type="video/mp4">
  Your browser does not support the video tag.
</video> 

## Figure out junctions in the game grid

Junctions in the game are where the character could change directions.

Ghosts (and the auto-driving pacman) should only change directions at a junction
where there is a choice in direcctions.  This will remove arbitrary reversals
in direction.

```{r}
# Movement map
roll_down  <- rbind(board[-1, ], rep(NA, 28)) 
roll_up    <- rbind(rep(NA, 28), board[-nrow(board), ]) 
roll_right <- cbind(board[,-1], rep(NA, 31))
roll_left  <- cbind(rep(NA, 31), board[,-ncol(board)])

move_left  <- board == '.' & roll_left  == '.'
move_right <- board == '.' & roll_right == '.'
move_up    <- board == '.' & roll_up    == '.'
move_down  <- board == '.' & roll_down  == '.'

moves <- list(
  left  = move_left ,
  right = move_right,
  up    = move_up   ,
  down  = move_down 
)

junction <- (move_left | move_right) & (move_up | move_down) #&
              # (move_left + move_right + move_up + move_down > 2)

mode(move_left)  <- 'integer'
mode(move_right) <- 'integer'
mode(move_up)    <- 'integer'
mode(move_down)  <- 'integer'
mode(junction)   <- 'integer'

# Plot junction locations where ghost/pacman movements may be changed
coords <- arrayInd(which(junction == 1), dim(junction))
junction_nr <- nr_duplicate(blank_board_nr)
nr_rect(junction_nr, coords[,2]*8 - 3, (32 - coords[,1])*8 - 3, 2, 2, 'white')
grid.raster(junction_nr, interpolate = FALSE)
```


## Movement engine


```{r eval=FALSE}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Open a a double-buffered x11() device.
# - turn antialiasing off (don't need it for pixel rendering)
# - inibit storage of a displaylist. not needed.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
x11(type = 'dbcairo', antialias = 'none')
dev.control(displaylist = 'inhibit')
  
  
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Given the row and column pick a random direction to move in
# Ensure that the direction chosen isn't the reverse of the current
# direction
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
choose_direction <- function(row, col, current) {
  dirs <- c('left', 'right', 'up', 'down')
  reverse <- c('right', 'left', 'down', 'up')
  remove <- which(reverse == current)
  idxs <- setdiff(1:4, remove)
  for (i in sample(idxs)) {
    if (moves[[i]][32 - row, col]) break
  }
  dirs[i]
}


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# 'pacman' state information
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
pacdir <- ""
dx     <- 0
dy     <- 0
row    <- 2
col    <- 2

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Current dots in the scene
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
cdots <- dots

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Ghost state information. This will be updated during the game
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
gh <- list(
  list(row =  2, col =  2, dx = 0, dy = 0, dir = "down"),
  list(row =  2, col = 27, dx = 0, dy = 0, dir = "down"),
  list(row = 30, col =  2, dx = 0, dy = 0, dir = "down"),
  list(row = 30, col = 27, dx = 0, dy = 0, dir = "down")
)
  
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create working buffer where the board will be rendered
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
board_nr <- nr_duplicate(blank_board_nr)
  

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
for (i in 1:100) {

  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Eat the dot where the pacman is
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  matches <- (cdots$y == row & cdots$x == col)
  cdots <- cdots[!matches,, drop = FALSE]
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # If pacman at a junction, then decide new direction
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  if (junction[32 - row, col]) {
    dir <- choose_direction(row, col, pacdir)
    pacdir <- dir
    if (dir == 'left' ) {dx = -1; dy =  0} else
    if (dir == 'right') {dx =  1; dy =  0} else
    if (dir == 'up'   ) {dx =  0; dy =  1} else
    if (dir == 'down' ) {dx =  0; dy = -1} else 
    {dx = 0; dy = 0}
  }
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # For each ghost:
  #  - if at a junction, choose a random new direction
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  for (i in seq_along(gh)) {
    if (junction[32 - gh[[i]]$row, gh[[i]]$col]) {
      dir <- choose_direction(gh[[i]]$row, gh[[i]]$col, gh[[i]]$dir)
      gh[[i]]$dir <- dir
      if (dir == 'left' ) {gh[[i]]$dx = -1; gh[[i]]$dy =  0} else
      if (dir == 'right') {gh[[i]]$dx =  1; gh[[i]]$dy =  0} else
      if (dir == 'up'   ) {gh[[i]]$dx =  0; gh[[i]]$dy =  1} else
      if (dir == 'down' ) {gh[[i]]$dx =  0; gh[[i]]$dy = -1} else 
      {gh[[i]]$dx = 0; gh[[i]]$dy = 0}
    }
  }
  
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Move pacman/ghosts from one location to the next in 8 steps
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  for (step in 1:8) {
    
    # Copy blank board into our working buffer
    nr_copy_into(board_nr, blank_board_nr)
    
    # Blit current dots into board
    nr_blit(board_nr, dot_nr, (cdots$x - 0.5) * 8, (cdots$y - 0.5) * 8)
    
    # Blit ghosts into board
    for (i in seq_along(gh)) {
      nr_blit(
        board_nr, 
        ghost[[i]][[gh[[i]]$dir]][[ bitwShiftR(step, 1L) %% 2 + 1L]], 
        x = gh[[i]]$col * 8 - 11 + step * gh[[i]]$dx, 
        y = gh[[i]]$row * 8 - 11 + step * gh[[i]]$dy
      )
    }
    
    # Blit pacman into board
    nr_blit(
      board_nr, 
      pacman[[pacdir]][[ bitwShiftR(step, 1L) %% 4 + 1L]], 
      x = col * 8 - 11 + step * dx, 
      y = row * 8 - 11 + step * dy
    )
    
    # Draw everything to screen
    dev.hold()
    grid.raster(board_nr)
    dev.flush()
    
    # regulate drawing speed. Replace with {eventloop} for actual game
    Sys.sleep(0.01)
  }
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Advance pacman location to new row/col
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  row <- row + dy
  col <- col + dx
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Advance ghosts to the next row/col
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  for (i in seq_along(gh)) {
    gh[[i]]$row <- gh[[i]]$row + gh[[i]]$dy
    gh[[i]]$col <- gh[[i]]$col + gh[[i]]$dx
  }
}
```

<video controls>
  <source src="./image/pacman-attractb.mp4" type="video/mp4">
  Your browser does not support the video tag.
</video> 














