---
title: "Triangles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Triangles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nara)
```


`{nara}` has three ways to draw triangles:

* `nr_polygon()`
* `nr_tri_mesh()`
* `nr_tri_coords()`

For single triangles, `nr_polygon()` is sufficient.  For large numbers of
polygons, the `nr_tri_*` functions will be much much faster.

This vignette demonstrates the drawing of two triangles using each of these 
functions.


## Demo setup

Define 5 points on the image which will be used as the vertices of two triangles


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Define image size 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
w <- h <- 300
th <- 16 # text height

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Define 5 points within the canvas
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
ax <- w/2     ; ay <- 0 + th
bx <- w/2     ; by <- h/2 + th     
cx <-   0     ; cy <- h/2 + th     
dx <- w/2     ; dy <- h            
ex <- w - th/2; ey <- h/2 + th

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Label the points in an image
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
nr <- nr_new(w, h)
nr_text_basic(nr, ax, ay, 'A', fontsize = th)
nr_text_basic(nr, bx, by, 'B', fontsize = th)
nr_text_basic(nr, cx, cy, 'C', fontsize = th)
nr_text_basic(nr, dx, dy, 'D', fontsize = th)
nr_text_basic(nr, ex, ey, 'E', fontsize = th)

plot(nr)
```



# Render triangles with `nr_polygon()`

Render two triangles by using two calls to `nr_polygon()`.

```{r echo = FALSE}
nr <- nr_new(w, h)

nr_polygon(nr, c(ax, bx, cx), c(ay, by, cy), fill = "blue")
nr_polygon(nr, c(bx, dx, ex), c(by, dy, cy), fill = "red")

plot(nr)
```


# Render triangles with `nr_tri_mesh()`

`nr_tri_mesh()` expects the triangles to be defined in two parts:

1. A matrix of vertex coordinates 
    *  `x` values in the first row, `y` values 
        in the second row.  (Further rows are ignored)
2. An integer matrix of indices.
    * The first 3 values in each column define the columns in the `vertices`
      matrix which define the triangle.
      
This mesh data definition is a similar format to the `mesh3d` class of data 
defined in the `{rgl}` package.

```{r}
# define all the vertex coordinates
vertices <- rbind(
  c(ax, bx, cx, dx, ex),
  c(ay, by, cy, dy, ey)
)

# Define the indices.  Each column in this matrix represents 1 triangle.
# The three numbers reprsent the column index in the `vertices` data
indices <- rbind(
  # 1  2
  c(1, 2),
  c(2, 4),
  c(3, 5)
)


nr <- nr_new(w, h)
nr_tri_mesh(nr, vertices, indices, c('blue', 'red'))

plot(nr)
```


# Render triangles with `nr_tri_coords()`

`nr_tri_coords()` requires just the raw coordinates of all vertices in a single
matrix.  The first row has the `x` coordinates, the second row has the `y`
coordinates.

Each group of 3 columns defines a triangle

```{r}
coords <- rbind(
  # <<<  1 >>>    <<<  2 >>>>
  c(ax, bx, cx,   bx, dx, ex),
  c(ay, by, cy,   by, dy, ey)
)

nr <- nr_new(w, h)
nr_tri_coords(nr, coords, color = c('blue', 'red'))

plot(nr)
```















